---
title: "Bayesian_comparison_DEGs"
author: "Cam Lunn"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Read in $\beta$ values and merge


```{r Using general (no grouping) LMM}

frequentist_LMM_results <- read.csv("R_data/LMM_results_general.csv") %>%
  dplyr::arrange(`Pr...t..`)

betas_individual <- read.csv("R_data/beta_filtered_genes_indv.csv") %>%
  rename(RTS_ID = gene)

betas_random <-  read.csv("R_data/beta_filtered_genes_random_20.csv") %>%
  rename(RTS_ID = gene)


overlap_individuals <- merge(betas_individual, frequentist_LMM_results, by = "RTS_ID") %>%
    arrange(Pr...t..)


overlap_individuals %>%
  filter(Pr...t.. <= 0.05)


#### Random
overlap_random <- merge(betas_random, frequentist_LMM_results, by = "RTS_ID") %>%
  arrange(Pr...t..)

# See the 39 overlapped:
overlap_random %>%
  filter(Pr...t.. <= 0.05)


```
When merging the random Beta values to the entire gene list used in frequentist analysis, the list of betas drops from 439 --> 395. This means our model identified 44 genes which were removed during QC steps. Thus, our model is most likely overfitting to noise since it is using ~10% of genes that were removed due to poor data quality.

When merging the individual Beta values, we go from 295 betas --> 273. This means our model using individual betas identified 22 genes which were removed during QC steps. 



```{r, fig.width = 11, fig.height = 7, fig.wide = TRUE, warning = FALSE, message = FALSE}

pval_cutoff = .05
library(ggVennDiagram)
library(dplyr)

gene_sets <- list(
  Condition1 = frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff) %>% 
    magrittr::use_series(RTS_ID),
  Condition2 = betas_random %>%
    magrittr::use_series(RTS_ID)
  # ,
  # Condition3 = betas_individual %>%
  #   magrittr::use_series(RTS_ID)
)

nrow(frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff))

random_beta_laplace_venn <- ggVennDiagram(
  gene_sets,
  category.names = c(paste("Frequentist DEGs:", nrow(frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff))) , paste("Betas Random:", nrow(betas_random))
                     # , "Betas Individual"
                     ),
  set_color = "black",
  label_color = "black", label_alpha = 1
) + scale_fill_gradient(low="beige",high = "red") + 
  labs(caption = "p < 0.05",
       title = "Overlap of DEGs using Laplace Priors") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
# +
#   theme(legend.position = "bottom")

ggsave("./images/random_beta_laplace_venn.pdf", plot = random_beta_laplace_venn, width = 11)


############# Individual

gene_sets <- list(
  Condition1 = frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff) %>% 
    magrittr::use_series(RTS_ID),
  # Condition2 = betas_random %>%
  #   magrittr::use_series(RTS_ID)
  # ,
  Condition2 = betas_individual %>%
    magrittr::use_series(RTS_ID)
)

individual_beta_laplace_venn <- ggVennDiagram(
  gene_sets,
  category.names = c(paste("Frequentist DEGs:", nrow(frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff))) , paste("Betas Individual:", nrow(betas_individual))
                     # , "Betas Individual"
                     ),
  set_color = "black",
  label_color = "black", label_alpha = 1
) + scale_fill_gradient(low="beige",high = "red") + 
  labs(caption = "p < 0.05",
       title = "Overlap of DEGs using Laplace Priors") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
# +
#   theme(legend.position = "bottom")

ggsave("./images/individual_beta_laplace_venn.pdf", plot = individual_beta_laplace_venn, width = 11)



```



###### Using Horseshoe


```{r}

frequentist_LMM_results <- read.csv("R_data/LMM_results_general.csv") %>%
  arrange(`Pr...t..`)

betas_individual_horse <- read.csv("R_data/beta_filtered_genes_indv_horseshoe.csv") %>%
  rename(RTS_ID = gene)
betas_individual_horse


overlap_individuals_horse <- merge(betas_individual_horse, frequentist_LMM_results, by = "RTS_ID")
overlap_individuals_horse %>%
  arrange(`Pr...t..`) %>%
  filter(`Pr...t..` < .05)



betas_random_horse <-  read.csv("R_data/beta_filtered_genes_rand_horseshoe.csv") %>%
  rename(RTS_ID = gene)
betas_random_horse

overlap_random_horse <- merge(betas_random_horse, frequentist_LMM_results, by = "RTS_ID")
overlap_random_horse %>%
  arrange(`Pr...t..`) %>%
  filter(`Pr...t..` < .05)



```


#### Venn d

```{r, fig.width = 11, fig.height = 7, fig.wide = TRUE, warning = FALSE, message = FALSE}

pval_cutoff = .05
library(ggVennDiagram)

gene_sets <- list(
  Condition1 = frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff) %>% 
    magrittr::use_series(RTS_ID),
  Condition2 = betas_random_horse %>%
    magrittr::use_series(RTS_ID)
  # ,
  # Condition3 = betas_individual %>%
  #   magrittr::use_series(RTS_ID)
)

nrow(frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff))

random_beta_horse_venn <- ggVennDiagram(
  gene_sets,
  category.names = c(paste("Frequentist DEGs:", nrow(frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff))) , paste("Betas Random:", nrow(betas_random_horse))
                     # , "Betas Individual"
                     ),
  set_color = "black",
  label_color = "black", label_alpha = 1
) + scale_fill_gradient(low="beige",high = "red") + 
  labs(caption = "p < 0.05",
       title = "Overlap of DEGs using Horseshoe Priors") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
# +
#   theme(legend.position = "bottom")

ggsave("./images/random_beta_horse_venn.pdf", plot = random_beta_horse_venn, width = 11)


############# Individual

gene_sets <- list(
  Condition1 = frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff) %>% 
    magrittr::use_series(RTS_ID),
  # Condition2 = betas_random %>%
  #   magrittr::use_series(RTS_ID)
  # ,
  Condition2 = betas_individual_horse %>%
    magrittr::use_series(RTS_ID)
)

individual_beta_horse_venn <- ggVennDiagram(
  gene_sets,
  category.names = c(paste("Frequentist DEGs:", nrow(frequentist_LMM_results %>%  
    filter(`Pr...t..`<= pval_cutoff))) , paste("Betas Individual:", nrow(betas_individual_horse))
                     # , "Betas Individual"
                     ),
  set_color = "black",
  label_color = "black", label_alpha = 1
) + scale_fill_gradient(low="beige",high = "red") + 
  labs(caption = "p < 0.05",
       title = "Overlap of DEGs using Horseshoe Priors") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
# +
#   theme(legend.position = "bottom")

ggsave("./images/individual_beta_horse_venn.pdf", plot = individual_beta_horse_venn, width = 11)



```

### Check overlap of all gene sets... did they find similar genes?


```{r, fig.width=11, fig.height=7}

overlap_individuals
overlap_individuals_horse
overlap_random
overlap_random_horse


gene_sets_full <- list(
  Condition1 = overlap_individuals %>% 
    magrittr::use_series(RTS_ID),
  Condition2 = overlap_individuals_horse %>%
    magrittr::use_series(RTS_ID),
  Condition3 = overlap_random %>%
    magrittr::use_series(RTS_ID),
  Condition4 = overlap_random_horse %>%
    magrittr::use_series(RTS_ID)
)
library(ggVennDiagram)
full_venn <- ggVennDiagram(
  gene_sets_full,
  category.names = c(paste("Laplace Betas Indiv:", nrow(overlap_individuals)) , 
                     paste("Hoseshoe Betas Indiv:", nrow(overlap_individuals_horse)),
                     paste("Laplace Betas Random:", nrow(overlap_random)),
                     paste("Horseshoe Betas Random:", nrow(overlap_random_horse))
                     ),
  set_color = "black",
  label_color = "black", label_alpha = 1
) + scale_fill_gradient(low="beige",high = "red") + 
  labs(caption = "p < 0.05",
       title = "Overlap of DEGs using ALL Priors") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))+
  scale_x_continuous(expand = expansion(mult = .2))


ggsave("./images/full_venn.pdf", plot = full_venn, width = 11)


```





```{r}

 full_overlap <- intersect(intersect(intersect(gene_sets_full$Condition1, gene_sets_full$Condition3), gene_sets_full$Condition4),  gene_sets_full$Condition2)

data.frame(full_overlap)

(frequentist_LMM_results %>%
  filter(Pr...t.. < .05))$RTS_ID

merge(data.frame(intersect(full_overlap, (frequentist_LMM_results %>%
  filter(Pr...t.. < .05))$RTS_ID)) %>% rename(RTS_ID = intersect.full_overlap...frequentist_LMM_results.....filter.Pr...t.....
), (frequentist_LMM_results %>%
  filter(Pr...t.. < .05)), by = "RTS_ID") %>%
  arrange(Estimate)

```

CDKN1C gene provides instructions for making a protein that helps regulate growth.
REN gene provides instructions for making a protein called renin, which is produced in the kidneys.
UMOD gene provides instructions for making a protein called uromodulin. This protein is produced by the kidneys and then excreted from the body in urine.
CALM2: Increased exosomal Calmodulin in the urine of patients with diabetic kidney disease (DKD) may be derived from either CALM1, CALM2
PTMAL expressed in kidney, associated with kidney diseases in mice





```{r, fig.width=11, fig.height=7}

betas_individual_horse
betas_random_horse
betas_individual
betas_random


gene_sets_full_full <- list(
  Condition1 = betas_individual %>% 
    magrittr::use_series(RTS_ID),
  Condition2 = betas_individual_horse %>%
    magrittr::use_series(RTS_ID),
  Condition3 = betas_random %>%
    magrittr::use_series(RTS_ID),
  Condition4 = betas_random_horse %>%
    magrittr::use_series(RTS_ID)
)


full_venn2 <- ggVennDiagram(
  gene_sets_full_full,
  category.names = c(paste("Laplace Betas Indiv:", nrow(betas_individual)) , 
                     paste("Hoseshoe Betas Indiv:", nrow(betas_individual_horse)),
                     paste("Laplace Betas Random:", nrow(betas_random)),
                     paste("Horseshoe Betas Random:", nrow(betas_random_horse))
                     ),
  set_color = "black",
  label_color = "black", label_alpha = 1
) + scale_fill_gradient(low="beige",high = "red") + 
  labs(caption = "p < 0.05",
       title = "Overlap of DEGs using ALL Priors") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))+
  scale_x_continuous(expand = expansion(mult = .2))


ggsave("./images/full_venn_complete.pdf", plot = full_venn2, width = 11)


```









